Unsupervised embeddings of organizations capture latent structure of scientific mobility

### 2020-04-12 paper/current/main.tex
reflect_sada's comment

Let $T_{ij}$ represents the total \textbf{flux}, the number of people who moved between organizations $i$ and $j$. $M_{i}$ and $M_{j}$ represent the size of an organization. According to our gravity model on the embedding space,

\begin{equation}
    \label{eq:gravity_model}
    \frac{T_{ij}}{M_{i} M_{j}} \propto \exp(\cos(v_i,v_j))
\end{equation}

where the cosine similarity between two vectors, $\cos(v_{i}, v_{j})$, is defined as,

\begin{equation}
    \cos(v_i, v_j) = \frac{v_i \cdot v_j}{\Vert v_j \Vert \Vert v_i  \Vert}
\end{equation}

In the case of scientific co-affiliation, all movements are without direction (symmetric), meaning that  $T_{ij} = T_{ji}$.
Let's denote $N$ as set of organizations in the data.
Then, the total flux to or from $i$ can be written as

\begin{equation}
	\label{eq:approxi_mass_term}
    \mu_{i} = \sum_{k \in N} T_{ik} \propto M_i
\end{equation}

It can be checked in Fig.\ref{fig:math_part}. With this approximation, we can rewrite Eq.~\ref{eq:gravity_model} as,
\begin{equation}
   \frac{T_{ij}}{\mu_{i} \mu_{j}} \propto \exp(\cos(v_i,v_j))
\end{equation}

@@

where the conditional probability of observing one word, $w_{0}$, given another, $w_{1}$, is defined as,

@@

where $v_w$ and $v'_w$ are ``input'' and ``output'' vector representations of each word in the vocabulary.  We use Skip-gram negative sampling, and following the work by Levy and Goldberg \cite{levy2014neural}, consider the model as implicitely performing matrix factorization as shown below,

\begin{equation}
    \label{eq:sgns_relation}
    \log \left(\frac{\#(w,c) |D|}{\#(w) \#(c)} \right) - \log k
\end{equation}

where $\#(w)\text{, } \#(w) \text{ and } \#(c)$ denotes the number of times word $w$, $c$ and word-context-pair ($w$, $c$) in the corpus respectively. $k$ is the number of negative samples and $|D|$ is the size of the corpus. Eq. \ref{eq:sgns_relation} is symmetry matrix, result vector representation  $v_w$ and $v'_w$ are identical. With these relations, for a specific pair of words $(w_i, w_j)$,

\begin{equation}
    \label{eq:sgns_relation_w,i, w_j}
    v_{w_i} \cdot v_{w_j} =  \log \left(\frac{\#(w_i, w_j) |D|}{\#(w_i) \#(w_j)} \right) - \log k
\end{equation}

Where $|D|$ and $k$ are constants, such that the functional form can be re-written as,

\begin{equation}
    \label{eq:sgns_relation_w,i, w_j}
     \frac{\#(w_i, w_j)}{\#(w_i) \#(w_j)}  \propto  \exp(v_{w_i} \cdot v_{w_j} )
\end{equation}

Combining Eq.~\ref{eq:approxi_mass_term}, we have

\begin{equation}
    \label{eq:sgns_relation_w,i, w_j_combining}
     \frac{T_{ij}}{\mu_{i} \mu_{j}}  \propto  \exp(v_{w_i} \cdot v_{w_j} )
\end{equation}

>>>>

According to our observation on the embedding space,

\begin{equation}
    \label{eq:gravity_model}
    \frac{T_{ij}}{M_{i} M_{j}} \propto -\exp(d_{ij}) = - \exp(1 - \cos(v_i,v_j)) \propto \exp(\cos(v_i,v_j)),
\end{equation}

where $T_{ij}$ represents the \textbf{flux}, the number of people who moved between organizations $i$ and $j$,  $M_{i}$ and $M_{j}$ represent the size of an organization,  $d_{ij} = 1 - \cos(v_{i}, v_{j})$ is the embedding distance, and $\cos(v_{i}, v_{j})=\frac{v_{i} \cdot v_{j}}{\lVert v_{i} \rVert \lVert v_{j} \rVert}$  is the cosine similarity between two vectors. In the case of scientific co-affiliation, we assume all movements are without direction (symmetric), meaning that  $T_{ij} = T_{ji}$.

To convert Eq. \ref{eq:gravity_model} to a flux-related term only, we replace  $M_*$ with $\mu_{i*}$, the total flux to or from $i$ which can be written as

\begin{equation}
	\label{eq:approxi_mass_term}
    \mu_{i} = \sum_{k \in N} T_{ik} \propto M_i,
\end{equation}
where $N$ as set of organizations in the data. It can be checked in Fig.\ref{fig:math_part}. With this approximation, we can rewrite Eq.~\ref{eq:gravity_model} as,
\begin{equation}
   \frac{T_{ij}}{\mu_{i} \mu_{j}} \propto \exp(\cos(v_i,v_j)).
\end{equation}

@@

where the conditional probability of observing one word, $w_{O}$, given another, $w_{I}$, is defined as,

@@

where $v_w$ and $v'_w$ are ``input'' and ``output'' vector representations of each word in the vocabulary.  We use Skip-gram negative sampling, and following the work by Levy and Goldberg \autocite{levy2014neural}, consider the model as implicitly performing matrix factorization as shown below,

\begin{equation}
    \label{eq:sgns_relation}
    \log \left(\frac{\#(w,c) |D|}{\#(w) \#(c)} \right) - \log k,
\end{equation}

where $\#(w)\text{, } \#(c) \text{ and } \#(w, c)$ denotes the number of times word $w$, $c$, and word-context-pair ($w$, $c$) in the corpus respectively. Variable $k$ is the number of negative samples and $|D|$ is the size of the corpus.

Mathematically, the maximization of the objective function of the model yields $V=L$ and $V'=R$ where L and R are the left and right singular vectors of the matrix Eq.\ref{eq:sgns_relation}, respectively. Note that $V=L=R=V'$ because Eq.\ref{eq:sgns_relation} is a real symmetric matrix. With these relations, for a specific pair of words $(w_i ,w_j )$,

\begin{equation}
    \label{eq:sgns_relation_w,i, w_j}
    v_{w_i} \cdot v_{w_j} =  \log \left(\frac{\#(w_i, w_j) |D|}{\#(w_i) \#(w_j)} \right) - \log k,
\end{equation}

where $|D|$ and $k$ are constants, such that the functional form can be re-written as,

\begin{equation}
    \label{eq:sgns_relation_w,i, w_j}
     \frac{\#(w_i, w_j)}{\#(w_i) \#(w_j)}  \propto  \exp(v_{w_i} \cdot v_{w_j} ).
\end{equation}

In our model, $\#(w_i)$ and $\#(w_j)$ are total flux $M_i$ and $M_j$ respectively,  and $\#(w_i,w_j)$ is just  $T_{ij}$, flux between $i$ and $j$. Then, we have the same expression with gravity model,

\begin{equation}
    \label{eq:sgns_relation_w,i, w_j_combining}
     \frac{T_{ij}}{\mu_{i} \mu_{j}}  \propto  \exp(v_{i} \cdot v_{j} ).
\end{equation}
